---
title: "STA100 Lab2"
author: "Chi Po Choi"
date: "10/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Quick start with R and RStudio

#### Working directory

Working directory is the default folder in which files are saved and read.

- In the **Files** tab, find/create a folder for your STA100 work. Click **More** then **Set As Working Directory**.

- Download the `GSK.csv` in [STA100B Piazza Resources ](https://piazza.com/ucdavis/fall2021/sta100b/resources)  and save it into your working directory. Then, run the codes `gsk_data = read.csv("GSK.csv")`

- Click on the Data `gsk_data` in Environment tab.

#### R scripts

- Click the "little green plus icon" and select **R Script**. Then you obtain a notebook that for your to put R codes. Type some codes, for example
  ```{r, eval=F}
  gsk_data = read.csv("GSK.csv", stringsAsFactors = TRUE)
  hist(gsk_data$sbp, main = "Hey")
  boxplot(sbp ~ exercise, data=gsk_data, main = "Hi")
  ```

- You can also **Run** line by line. Put your cursor to the line your want to run and click **Run**. You can also use hot-key **Command+Enter** in mac (or **Control+Enter** in Windows?).

- Click the floppy icon to save your R script in your working directory. Usually we save R scripts with file extension `.R`, for example, `lab2_is_the_best.R`.

# Exercise in `Discussion_02.pdf`

Get the `Discussion_02.pdf` prepared by Maxime in [STA100B Piazza Resources ](https://piazza.com/ucdavis/fall2021/sta100b/resources) . 

#### Book Problems

Let's work on some the problems during the Lab.

- Q1
  12:20pm-12:30pm

- Q2
  12:40pm-12:50pm

####  R Portion

Online you will find the file “GSK.csv”. The csv file has the following columns:
```
Column 1. sysbp: The systolic blood pressure of the subject (mmHg).
Column 2. gender: The gender, with levels F and M.
Column 3. married: Y if the subject was married, N if not.
Column 4. exercise: With levels L = low, M = medium, H = high.
Column 5. age: The age of the subject in years.
Column 6. stress: With levels LS = low, MS = medium, HS = high.
Column 7. educatn: With levels LE = low, ME = medium, HE = high.
```
(a) Create a histogram of systolic blood pressure. Which interval has the highest number of subjects in it?
(b) Create a boxplot of systolic blood pressure by exercise group. Which group has the highest median?
(c) Create a barplot for the stress levels. Which group has the least subjects in it?
(d) Create a mosaic plot between stress and education. Of those who have “medium” education, what is the most
common stress level?

```{r, eval=F}
gsk_data = read.csv("GSK.csv", stringsAsFactors = TRUE)

hist(gsk_data$sbp, main = "Hey")

boxplot(sbp ~ exercise, data=gsk_data, main = "Hi")

plot(gsk_data$stress, main = "Yo")

mosaicplot(~ stress + educatn, data = gsk_data, shade = TRUE)
```

