---
title: "STA100 Lab7"
author: "Chi Po Choi"
date: "11/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise in `Discussion-07.pdf`

Get the `Discussion-07.pdf` prepared by Maxime in [STA100B Piazza Resources ](https://piazza.com/ucdavis/fall2021/sta100b/resources) . 

# Hints

#### Discussion-07

- Q2 (a-b)

  | Brand |   A  |   B  |  C  |
  |:-----:|:----:|:----:|:---:|
  |   $\bar{y}_i$   | 10.5 | 10.6 | 8.2 |
  |   $s_i$   |  2.1 |  2.2 | 4.8 |
  |   $n_i$   |  25  |  25  |  25 |


  $$
  \begin{aligned}
  H_0:& \mu_A = \mu_B = \mu_C \\
  H_A:& \text{Not all $\mu_A, \mu_B, \mu_C$ are equal.}
  \end{aligned}
  $$

  The F test statistics
  $$
  \begin{aligned}
  \bar{\bar{y}}
  &=
  \frac{
  25 \times 10.5 + 25 \times 10.6 + 25 \times 8.2
  }{
  25+25+25
  }
  =
  9.7667
  \\
  MS(\text{Between})
  &=
  \frac{
    SS(\text{Between})
  }{
    \text{d.f.}(\text{Between})
  }
  =
  \frac{
    25 \times (10.5 - 9.7667)^2
  + 25 \times (10.6 - 9.7667)^2
  + 25 \times (8.2 - 9.7667)^2
  }{
    3 - 1
  }
  =
  46.0833
  \\
  MS(\text{Within})
  &=
  \frac{
    SS(\text{Within})
  }{
    \text{d.f.}(\text{Within})
  }
  = 10.6159 \quad \text{(given in this question)}
  \\
  F
  &=
  \frac{MS(\text{Between})}{MS(\text{Within})}
  =
  \frac{46.0833}{10.6159}
  =
  4.3410
  \end{aligned}
  $$

  Look at the F distribution table:
  
  - Numerator df $= 3 - 1 = 2$
  - Denom. df $= 75 - 3 = 72$
  - F $\sim (4.18, 4.98)$
  - p-value $\sim (0.01, 0.02)$

- Q2 (c)
  
  $\alpha = 0.05$, while p-value is between $0.01$ and $0.02$.
  
  p-value is smaller than $\alpha$, therefore $H_0$ null hypothesis is rejected.

- Q2 (d)
  
  The ANOVA analysis here assumes equal variance across all brand (though they could have difference population mean).
  The sample standard deviation of Brand C is a much higher than that of A and B. It is suspicious that Brand C could have differerent variance than A and B.


#### Homework 5

- R Homework I

Lets load the DRP data and plot it. Note that we use the function `factor` to set up the categorical variable `Treatment` in the way that "Treated" is at front and "Control" is at back.
```{r}
data_drp <- read.csv("DRP.csv")
data_drp$Treatment = factor(data_drp$Treatment, levels = c("Treated", "Control")) #important line!
boxplot(Response ~ Treatment, data = data_drp)
```


Lets compute the two-sided confidence interval for the difference in population means. The difference is
$$
\mu_\text{Treatment} - \mu_\text{Control}
$$
We use the function `t.test` to compute the two-sided confidence interval:
```{r}
t.test(Response ~ Treatment,
       data = data_drp,
       conf.level = 0.95)
```

From the output of `t.test`, we see that
the 95 percent confidence interval of the difference in average DRP score between  "Treated" and "Control" is $(1.23302, 18.67588)$.


We use the function `t.test` again, but with the option `alternative = "greater"` for the hypothesis
$$
\begin{aligned}
H_0:& \quad \mu_\text{Treatment} - \mu_\text{Control} \leq 0 \\
H_A:& \quad \mu_\text{Treatment} - \mu_\text{Control} > 0
\end{aligned}
$$
```{r}
t.test(Response ~ Treatment,
       data = data_drp,
       alternative = "greater")
```
We can see the test statistics and p-value for this hypothesis.


- R Homework II

Lets load the IQ data and plot it.
```{r}
data_iq <- read.csv("IQ.csv")
data_iq$group <- factor(data_iq$group, levels = c("A", "B", "C"))

boxplot(iq ~ group, data = data_iq)
```


We may consider the hypothesis 
$$
\begin{aligned}
H_0:& \quad \mu_A = \mu_B = \mu_C \\
H_A:& \quad \text{not all equal}.
\end{aligned}
$$
Let use the `lm` and `anova` functions to find out the F test corresponding to this hypothesis. 
```{r}
model_iq <- lm(iq ~ group, data = data_iq)
anova(model_iq)
```

The F test statistics is 20.016 with d.f. being 2.


We can plot the Normal Probability Plot for IQ scores, with all group together.
```{r}
qqnorm(data_iq$iq, main = "Normal Probability Plot for IQ scores")
qqline(data_iq$iq)
```

We can plot the Normal Probability Plot for IQ scores with each group individually.
```{r}
index_A = data_iq$group == "A"
qqnorm(data_iq$iq[index_A], main = "Normal Probability Plot for IQ scores in group A")
qqline(data_iq$iq[index_A])

index_B = data_iq$group == "B"
qqnorm(data_iq$iq[index_B], main = "Normal Probability Plot for IQ scores in group B")
qqline(data_iq$iq[index_B])

index_C = data_iq$group == "C"
qqnorm(data_iq$iq[index_C], main = "Normal Probability Plot for IQ scores in group C")
qqline(data_iq$iq[index_C])
```
